---
title: "Digit Classification"
author: "Rishabh"
date: "6 February 2015"
output:
  html_document:
    toc: yes
---

```{r, Global Options and Load Libraries, echo=FALSE, include=FALSE}
# some useful global defaults
if(!require("knitr")) install.packages("knitr")
opts_chunk$set(fig.width=8, fig.height=5, warning=FALSE, message=FALSE, include=TRUE, echo=TRUE, cache=TRUE, cache.comments=FALSE)

if(!require("mvtnorm")) install.packages("mvtnorm")
if(!require("ggplot2")) install.packages("ggplot2")
if(!require("dplyr")) install.packages("dplyr")
if(!require("doParallel")) install.packages("doParallel")

#setwd("c:/Users/Rishabh/Documents/BGSE Material/Sem2/14D005 Machine Learning/Seminar2/HW/")
```

#Digit Classification

##Sourced Files
```{r, Source Helper Files and import datasets}
source("KNN_Classification.R")

#Training Data
train.data <- read.csv("training.csv", header = F)
trueClasses <- train.data[1:10,1]
train.data <- train.data[1:10,-1]

#Test Data
test.data <- read.csv("test.csv", header = F)

```


##Train Classifier
###Calculate Distance matrix
```{r, Calculate Distance matrix}
#detect cores for parallel computing
cores <- detectCores()

#Calculate Distance Matrix
cl <- makeCluster(cores)
registerDoParallel(cl)
clusterExport(cl, c("train.data", "calc_dist"))

dist_matrix <- parApply(cl, train.data, 1, function(row) (calc_dist(row, train.data)))

stopCluster(cl)

```

###Calculate nearest neighbours matrix
```{r, Calculate nearest neibours matrix}

#Calculate the neighbours
cl <- makeCluster(cores)
registerDoParallel(cl)
clusterExport(cl, c("dist_matrix"))
system.time(orderMat <- as.matrix(t(parApply(cl, dist_matrix, 1, order))))
stopCluster(cl)

```

###Classify for k = 1 and calculate empirical error and accuracy
```{r, Classify for k = 1}
k <- 1
k.orderMat <- as.matrix(orderMat[,2:(k+1)])

#Classify
cl <- makeCluster(cores)
registerDoParallel(cl)
clusterExport(cl, c("k.orderMat", "trueClasses"))
system.time(predictedClass <- as.character(parApply(cl, k.orderMat, 1, 
                                     function(x) names(which.max(table(trueClasses[x]))))))
stopCluster(cl)
#accuracy

accuracy <- mean(predictedClass == trueClasses)
error_rate <- 1 - accuracy

(round(c(accuracy = accuracy, "empirical error" = error_rate) * 100, 3))
```

###Selecting the best k
Calculate error rates for different k and select the one with the minimun error.
Although this is likely to be overfitted as there is no k fold validation done,
but due to lack of computing power and sub optimal functions, I proceed with it.
```{r, Selecting the best k}
#selecting the best k
k_seq <- seq.int(from = 1, to = 51, by = 2)
k_seq <- c(k_seq, 71, 91, 111, 131, 151, 171)

cl <- makeCluster(cores)
registerDoParallel(cl)

error_rate <- rep(NA, length(k_seq))

for(i in 1:length(k_seq)) {
  k.orderMat <- as.matrix(orderMat[,2:(i+1)])
  clusterExport(cl, c("k.orderMat", "trueClasses"))
  predictedClass <- as.character(parApply(cl, k.orderMat, 1, 
                                          function(x) names(which.max(table(trueClasses[x])))))
  error_rate[i] <- 1 - mean(predictedClass == trueClasses)
}
stopCluster(cl)
error_rate <- data.frame(k = k_seq, error = error_rate)

#plot the error rates
ggplot(data = error_rate, aes(x = k, y = error)) +
  geom_point() +
  geom_line() +
  xlab("k -Number of nearest neighbors") +
  ylab("Misclassification error") +
  theme_bw(base_size = 14) +
  scale_colour_brewer(palette="Set1")

k <- error_rate[which.min(error_rate[,2]),1]
(c("Best k" = k))
```


##Test data on the selected k
```{r, Test data}
#Calculate Distance
 
#Calculate the neighbours
cl <- makeCluster(cores)
registerDoParallel(cl)
clusterExport(cl, c("dist_matrix"))
system.time(orderMat <- as.matrix(t(parApply(cl, dist_matrix, 1, order))[,1:k]))
stopCluster(cl)

#Classify
cl <- makeCluster(cores)
registerDoParallel(cl)
clusterExport(cl, c("orderMat", "trueClasses"))
system.time(predictedClass <- as.character(
  parApply(cl, orderMat, 1, function(x) names(which.max(table(trueClasses[x]))))))
stopCluster(cl)

#Predicted Class Table
table(predictedClass)

```

##Write predictions to csv
```{r, write predictions to csv, eval = T}
#write to csv
write.csv(x = predictedClass, file = "predictedDigitsNew.csv")

```

